What can GitHub and Googlecode Issues tell us about Exceptions in Android?

### 2014-08-29 android-stacks.tex
changes along the paper, intro

This paper reports a repository mining study which analyzed all the issues created on
 Android projects available in GitHub (482 applications) and Google code (141 appplications)
Overall, our research set includes XXXX issues from which XXX stack traces 
were extracted. Such stack traces were distilled and analyzed in 
combination with source code and bytecode analysis. In this study some patterns of exception
(mis)-use were consistently detected such as: a high prevalence of NullPointerExceptions and 
other java.lang exceptions reported on issues - they were found in aproximately 50\% of all analyzed 
exception stack traces; undocumented runtime exceptions signaled by third party code - which is a serious threat to program
dependability; unexpected wrappings (e.g., Errors being wrapped in checked exceptions), revealing that Java hybrid exception
model is not fully used according to its purpose.

@@

Along the recent years, we have witnessed an astonishing increase in the number 
mobile applications publicly available. Such increase is motivated by pervasiveness of the World Wide Web 
and the unprecedent increase of global smartphone shipments -
the mobile industry shiped over 315 million smartphones in the last quarter of 2013 [ref-IO].

@@

In this context, Android has become the leading plaftorm of such applictions - thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus its
 own device) [ref].  Google estimates that there is over 1 billion Android active users per month [ref-IO],
and the number of apps built on top of Android increases in a daily rate.

@@

If on the one hand such applications extend the phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to cope to new challenges imposed by mobile environmet.
Such applications have to cope with an increasing number of abnormal
computation states, arising from: programmming errors;
 noisy user inputs; faults in underlying middleware or hardware; 
faults on third party services and libraries, compatibility issues [ref], 
conection with devices e.g., as GPS, camera, memory and batery restrictions
among others.

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps.
% as any other modern system~\cite{bruntink2006discovering}. 
The Android platform uses the exception handling mechanism embeeded
 in Java to deal with such conditions. Although the exception handling mechanism
~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
As a consequence that the exception handling code may become ilself a source of faults, 
and may affect the system the other way around, making the system even less robust ~\cite{coelho2011unveiling}[ref FU CHEN].

Google estimates that Android has over 1 billion active users per month, 
who pull out theirs phones and check it over 100 bilion times. 
In the same rate the number of mobile users increase, also increases 
the number of users affected by faulty apps and app crashes.

@@

Our conjecture is that several Android applicatons may be suffering from similar 
vulnerabilities on the exception handling code of the applications themselves or 
on the underlaying platform, and that such vulnerabiliries may be leading to
 application crashes -  as Uncaught Exceptions lead to app to crashes in
Android and is one of the main causes of application crashes in Java [ref AST] .

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, or log file [REF-STACK-TRACE-CRASH]. 
A stack trace includes a detail message and the execution stack frame when 
the exception occurred. A typical Java stack trace consists of an ordered list of 
methods that were active on the call stack before the exception has occurred [REF-STACKTRACE-MANUAL].

@@

Due to the high demand for responsiveness in mobile environment (mobile apps can easily become 
unresponsivedue due to resurce limitation and excessive network access [shimidit and arie]) the
Android Application Framework also provides APIs to support concurrent developement (e.g. AsyncTask).
As this framework is developed in Java, it reuses the embedded Java Exception model 
to detect and handle abnormal computation states.

@@

When an exception is thrown, it causes a transfer of control from the point
where the exception occurred to a point that can be specified by the programmer
(exception handler) - in Java it is represented by the try-catch block . A common way of  
propagating exceptions in Java programs is through exception wrapping
 (also known as exception chaining). Exception wrapping allows one exception 
to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} presents 
a stack trace associated which illustrates an exception wrapping. The bottom 
part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

@@

>>>>

This paper reports a repository mining study which analyzed all the issues created on
Android projects available in GitHub (482 applications) and Google code (141 appplications)
Overall, our research set includes XXXX issues from which XXX stack traces 
were extracted. Such stack traces were distilled and analyzed in 
combination with source code and bytecode analysis. This study 
consistently identifies common characteristics of reported 
exception stacks traces - e.g., the NullPointerException was the reported cause 
on 30,08\% of all analyzed exception stack traces, and ewas mentioned in approximatelly 
50% of all projects analyzed. Moreover,  the cross-project analysis 
of exception stack traces enabled to consistenly reveal   
fragilities of exception handling code: 
(1) undocumented runtime exceptions signaled by third party code,
(2) undocumente checked exceptions thrown by Android native code
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
Such fragilities impose challenges to application robusness not only 
in Android environemt but also to other systems based on Java exception model.

@@

Along the recent years, we have witnessed an astonishing increase in the number 
mobile applications being developed and some of them being largely used. 
Such increase is motivated by pervasiveness of the World Wide Web 
and the unprecedent increase of global smartphone shipments -
the mobile industry shiped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.

@@

In this context, Android has become a leading plaftorm for such applictions - thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
 own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
and the number of apps built on top of Android increases in a daily rate.

%ref: http://www.gartner.com/newsroom/id/2645115
% Gartner, Inc. Press release, 2012. www.gartner.com/newsroom/id/2237315.
% http://www.gartner.com/newsroom/id/2645115

@@

In the same rate the number of a users increase, also increases 
the number of users affected by application faults and crashes.
If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face new threats to application robustness imposed by mobile environmet.
Such applications have to cope with an increasing number of abnormal
computation states, arising from: programmming errors;
 noisy user inputs; faults in underlying middleware or hardware; 
faults on third party services and libraries, compatibility issues [ref], 
conection with devices e.g., as GPS, camera, memory and batery restrictions
among others.

%Hence, in the same rate the number of mobile users increase, also increases 
%the number of users affected by such threats.

%faulty apps and app crashes.

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps.
% as any other modern system~\cite{bruntink2006discovering}. 
The Android platform uses the exception handling mechanism embeeded
 in Java to deal with such conditions. Although the exception handling mechanism
~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
As a consequence that the exception handling code may become ilself a source of faults, 
and may affect the system the other way around, making the system even less robust ~\cite{robillard2003static,coelho2011unveiling}.

% Google estimates that Android has over 1 billion active users per month, 
% who pull out theirs phones and check it over 100 bilion times.

@@

Our conjecture is that several Android applicatons may be suffering from similar 
fragilities on the exception handling code (of the applications themselves or 
the underlaying platform), and that such fragilities may be leading to
decrease the application robustness, leading for instance to the Uncaught Exception problem
 are one of the main causes of application crashes ~\cite{jo2004uncaught} ~\cite{zhang2012amplifying}.

%[ref AST].

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, 
or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of an ordered list of 
methods that were active on the call stack before the exception has occurred,  
 and the fully qualified name of such exception.

@@

Due to the high demand for responsiveness in mobile environment (mobile apps can easily become 
unresponsivedue due to resurce limitation and excessive network access [shimidit and arie]) the
Android Application Framework also provides APIs to support concurrent developement (e.g. AsyncTask).

Android platform is mostly implemented in Java, uses JNI[ref] to access native C code, and 
reuses the embedded Java Exception model to detect and handle abnormal computation states.

@@

%When an exception is thrown, it causes a transfer of control from the point
%where the exception occurred to a point that can be specified by the programmer
%(exception handler) - in Java it is represented by the try-catch block .

In Java, once an exception is thrown, the JVM looks for the nearest enclosing exception handler
(in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
 To look for the handler on the invocation stack is claimed to increase software reusability, 
since the invoker of an operation can handle it in a wider context [ref-miller].

INCLUIR REF:
Miller, R.; Tripathi, A.: Issues with Exception Handling in Object-Oriented Systems. In:
ECOOP’97. LNCS, vol. 1241, Springer (1997) 85–103.

 A common way of  propagating exceptions in Java programs is through exception wrapping
 (also known as exception chaining). Exception wrapping allows one exception 
to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} presents 
a stack trace associated which illustrates an exception wrapping. The bottom 
part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

(1)

@@

\begin{table}
\centering
\begin{tabular}{ll}
 \bfseries{Method} & \bfseries{Project} & & \bfseries{Ocurrences}\\
    \hline
android.os.AsyncTask &	122 & 549  \\
android.view.LayoutInflater & 63 & 163 \\
android.support.v4.* & 100 & 512\\
\hline
  \end{tabular}
\caption{Elements tresponsible for the wrapping.}
\label{tab:wrapping01}
\end{table}
